Spectral Experts for Estimating Mixtures of Linear Regressions

### 2013-06-14 doc/icml-presentation/presentation.tex
Included a note about assumptions and the rest as usual

  \begin{tikzpicture}
    % x, y
    \llhood{0}{0};
    \node[scale=0.7] at (mle) {x};
    \node<2-> at ($(mle) + (0.6cm,0)$) {$\mathmb{\theta^*_{\textrm{MLE}}}$};
    \node[scale=0.7] at (em1) {x};
    \node<3-> at ($(em1) + (0.5cm,0)$) {$\mathmr{\hat\theta_{\textrm{EM}}}$};
    \node[scale=0.7] at (em2) {x};
    \node<3-> at ($(em2) + (0.5cm,0)$) {$\mathmr{\hat\theta_{\textrm{EM}}}$};

    \draw<4->[latex-latex,DarkGreen,line width=1pt] ($(mle) + (-0.8cm,0.8cm)$) -- node[above]{$\mathmg{\epsilon}$} ($(mle) + (+0.8cm,0.8cm)$);
  \end{tikzpicture}

@@

\begin{frame}
  \frametitle{Related Work}
  \begin{itemize}
    \item<1-> Method of Moments [Pearson, 1894]
    \item<2-> Observable operator models
    \begin{itemize}
      \item Control Theory [Ljung, 1987]
      \item Observable operator models [Jaeger, 2000; Littman / Sutton / Singh, 2004]
      \item \alert<2>{Hidden Markov models [Hsu/Kakade/Zhang, 2009]}
      \item Low-treewidth graphs [Parikh et al., 2012]
      \item Weighted finite state automata [Balle \& Mohri, 2012]
    \end{itemize}
     \item<3-> Parameter Estimation
  \begin{itemize}
    \item Mixture of Gaussians [Kalai/Moitra/Valiant, 2010]
    \item \alert{Mixture models, HMMs [Anandkumar/Hsu/Kakade, 2012]}
    \item Latent Dirichlet Allocation [Anandkumar/Hsu/Kakade, 2012]
    \item Stochastic block models [Anandkumar/Ge/Hsu/Kakade, 2012]
    \item Linear Bayesian networks [Anandkumar/Hsu/Javanmard/Kakade, 2012]
  \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Aside: Tensor Operations}
  \splitcolumn{%
    \begin{itemize}
      \item \tikzmark{tensoring} Tensoring
        \begin{align*}
          x\tp{3}       &= x \otimes x \otimes x \\
          x\tp{3}_{ijk} &= x_i x_j x_k
        \end{align*}
      \item<2-> \tikzmark{innerp}Inner product
        \begin{align*}
          \innerp{A}{B} &= \sum_{ijk} A_{ijk} B_{ijk} \\
          \uncover<3->{%
            &= \innerp{\vvec{A}}{\vvec{B}} 
          }
        \end{align*}
    \end{itemize}
  }{%
  }
  \begin{canvas}
    \point{mark}{(1.0cm,0)};
    \point{tensoring-pos}{({pic cs:tensoring} -| mark)}
    % Tensoring
    \tikzcube{tensoring}{black,fill=white}{($(tensoring-pos) + (0,-3cm) + (0,0,0)$)}{1}{1}{1};
    \node at ($(tensoring) + (1.0cm,-0.3cm)$) {$=$};
    \tikzcube{v1}{black,fill=white}{($(tensoring) + (2.5cm,-0.3cm)$)}{1}{0.3}{0.3};
    \node (lbl1) at ($(v1) + (0.40cm,-0.1cm)$) {$\times$};
    \tikzcube{v2}{black,fill=white}{($(v1) + (1cm,0.3cm)$)}{0.3}{1}{0.3};
    \node at ($(lbl1) + (1cm,0.0cm)$) {$\times$};
    \tikzcube{v3}{black,fill=white}{($(v2) + (1cm,-0.3cm)$)}{0.3}{0.3}{1};

@@

    % x, y
    \node[anchor=west,scale=1.0] (params) at (1.5cm,0) {%
      $\begin{bmatrix} \pi_1 \\ \pi_2 \\ \vdots \\ \pi_k \end{bmatrix}  
       \begin{bmatrix} 
                 &         &       &         \\
                 &         &       &         \\
         \beta_1 & \beta_2 & \dots & \beta_k \\
                 &         &       &         \\
                 &         &       &         
       \end{bmatrix} $
      };
    %\node[anchor=west,right=0.1cm of params] {%
    %\includegraphics[width=4cm,height=2cm,keepaspectratio]{figures/disc.png}
    %};

@@

  \begin{canvas}
    \input{spectral-experts.tikz}
    \node[style=txt] at (reg-label) {\alert<2>{regression}};
    \node[style=txt] at (tf-label) {\alert<3>{tensor factorization}};
  \end{canvas}
\end{frame}

@@

    \begin{align*}
      \robustaltm<-2>{\hat M_2}{\mathmb{\hat M_3}} &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} \left(y^2 - 
      \robustaltm<-2>{\innerp{M}{x\tp{2}}}{\mathmb{\innerp{M}{x\tp{3}}}}
      - \textrm{bias}\right)^2 
      \uncover<2->{+
      \robustaltm<2>{\mathmb{\ub{\|M\|_{*}}_{\sum_i \sigma_i(M)}}}
        {\|M\|_{*}}
      } \\
    \end{align*}

@@

  \begin{canvas}
    \input{spectral-experts.tikz}
    \node[style=txt] at (reg-label) {\robustalt<1>{\color{blue}\textbf{low-rank}}{low-rank}regression};
    \node[style=txt] at (tf-label) {tensor factorization};

@@

  \begin{canvas}
    % x, y
    \node[anchor=west] (em) at (-6.5cm,-0.0cm) {%
    \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1833.pdf}
      \includegraphics<2-3>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-1833.pdf}
      \includegraphics<4>[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-1833.pdf}
      \includegraphics<5->[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-Spectral+EM-1833.pdf}
      };

@@

    \draw<3>[latex-latex,DarkGreen,line width=1pt,dashed] ($(mle) + (-1.2cm,0.8cm)$) -- node[above]{$\mathmg{\epsilon}$} ($(mle) + (+1.2cm,0.8cm)$);
    \node<4->[scale=0.7] at (spec) {x};
    \node<4-> at ($(spec) + (0.5cm,0.3cm)$) {$\mathmg{\hat\theta_{\textrm{spec}}}$};
    \draw<4->[dashed,DarkGreen,line width=0.7pt] ($(spec)-(0.5cm,0)$) -- ($(spec)+(3.5cm,0)$);

    \draw<5->[-latex,smooth,line width=1pt,DarkGreen] ($(spec) + (-0.1cm,+0.05cm)$) -- ($(mle) + (-0.30cm,+0.10cm)$);
    \node<5->[scale=0.7] at (mle) {x};
    \node<5->[anchor=west] at ($(mle) + (0.3cm,0)$) {$\mathmb{\hat\theta}_{\textrm{spec + EM}}$};
    \draw<5->[dashed,blue,line width=0.7pt] ($(mle)-(0.4cm,0)$) -- ($(mle)+(0.4cm,0)$);
  \end{tikzpicture}

@@

\begin{frame}
  \frametitle{Conclusions}
  \begin{itemize}
    \item<1-> Expose tensor factorization structure through regression.
    \item<2-> Consistent estimator for the mixture of linear regressions with polynomial sample and computational complexity.
    \item<3-> Empirically, method of moment estimates can be a good initialization for EM.
    \item<4-> How can we handle other discriminative models?
      \begin{itemize}
          \item<5-> Dependencies between $h$ and $x$ (mixture of experts).
          \item<6-> Non-linear link functions (logistic regression).
      \end{itemize}
  \end{itemize}
\end{frame}

@@

% TikZ
%\begin{canvas}
%  \node (img1) at (0,0) {\includegraphics[height=0.6\textheight,width=0.4\linewidth,keepaspectratio]{}};
%\end{canvas}

% Notes 
% \note[item]{}

% 2-column
% \begin{columns}
%   \begin{column}{0.48\textwidth}
%     \begin{itemize}
%         \item
%     \end{itemize}
%   \end{column}
%   \hfill
%   \begin{column}{0.48\textwidth}
%     \begin{itemize}
%         \item
%     \end{itemize}
%   \end{column}
% \end{columns}

% TikZ
%\begin{canvas}
%  \node (img1) at (0,0) {\includegraphics[height=0.6\textheight,width=0.4\linewidth,keepaspectratio]{}};
%\end{canvas}

% Notes 
% \note[item]{}

% 2-column
% \begin{columns}
%   \begin{column}{0.48\textwidth}
%     \begin{itemize}
%         \item
%     \end{itemize}
%   \end{column}
%   \hfill
%   \begin{column}{0.48\textwidth}
%     \begin{itemize}
%         \item
%     \end{itemize}
%   \end{column}
% \end{columns}

>>>>

  \begin{tikzpicture}
    % x, y
    \llhood{0}{0};
    \node<2->[scale=0.3,circle,fill=black] at (mle) {};
    \node<2-> at ($(mle) + (0.6cm,0)$) {$\mathmb{\theta_{\textrm{MLE}}}$};
    \node<3->[scale=0.3,circle,fill=black] at (em1) {};
    \node<3-> at ($(em1) + (0.5cm,0)$) {$\mathmr{\theta_{\textrm{EM}}}$};
    \node<3->[scale=0.3,circle,fill=black] at (em2) {};
    \node<3-> at ($(em2) + (0.5cm,0)$) {$\mathmr{\theta_{\textrm{EM}}}$};

    %\draw<4->[latex-latex,DarkGreen,line width=1pt] ($(mle) + (-0.8cm,0.8cm)$) -- node[above]{$\mathmg{\epsilon}$} ($(mle) + (+0.8cm,0.8cm)$);
  \end{tikzpicture}

@@

\begin{frame}
  \frametitle{Related Work}
  \begin{itemize}
    \item<1-> Method of Moments [Pearson, 1894]
    \item<2-> Observable operators
    \begin{itemize}
      \item Control Theory [Ljung, 1987]
      \item Observable operator models [Jaeger, 2000; {\small{Littman/Sutton/Singh, 2004}}]
      \item Hidden Markov models [Hsu/Kakade/Zhang, 2009]
      \item Low-treewidth graphs [Parikh et al., 2012]
      \item Weighted finite state automata [Balle \& Mohri, 2012]
    \end{itemize}
     \item<3-> Parameter Estimation
  \begin{itemize}
    \item Mixture of Gaussians [Kalai/Moitra/Valiant, 2010]
    \item \alert{Mixture models, HMMs [Anandkumar/Hsu/Kakade, 2012]}
    \item Latent Dirichlet Allocation [Anandkumar/Hsu/Kakade, 2012]
    \item Stochastic block models [Anandkumar/Ge/Hsu/Kakade, 2012]
    \item Linear Bayesian networks [Anandkumar/Hsu/Javanmard/Kakade, 2012]
  \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Aside: Tensor Operations}
  \splitcolumn{%
    \begin{itemize}
      \item \tikzmark{tensoring} Tensor Product
        \begin{align*}
          x\tp{3}       &= x \otimes x \otimes x \\
          x\tp{3}_{ijk} &= x_i x_j x_k
        \end{align*}
      \item<2-> \tikzmark{innerp}Inner product
        \begin{align*}
          \innerp{A}{B} &= \sum_{ijk} A_{ijk} B_{ijk} \\
          \uncover<3->{%
            &= \innerp{\vvec{A}}{\vvec{B}} 
          }
        \end{align*}
    \end{itemize}
  }{%
  }
  \begin{canvas}
    \point{mark}{(1.0cm,0)};
    \point{tensoring-pos}{({pic cs:tensoring} -| mark)}
    % Tensoring
    \tikzcube{tensoring}{black,fill=white}{($(tensoring-pos) + (0,-3cm) + (0,0,0)$)}{1}{1}{1};
    \node at ($(tensoring) + (1.0cm,-0.3cm)$) {$=$};
    \tikzcube{v1}{black,fill=white}{($(tensoring) + (2.5cm,-0.3cm)$)}{1}{0.3}{0.3};
    \node (lbl1) at ($(v1) + (0.40cm,-0.1cm)$) {$\times$};
    \tikzcube{v2}{black,fill=white}{($(v1) + (1cm,0.3cm)$)}{0.3}{1}{0.3};
    \node at ($(lbl1) + (1cm,0.0cm)$) {$\times$};
    \tikzcube{v3}{black,fill=white}{($(v2) + (1cm,-0.3cm)$)}{0.3}{0.3}{1};

@@

    % x, y
    \node[anchor=west,scale=1.0] (params) at (1.5cm,0) {%
      $\begin{bmatrix} \pi_1 \\ \pi_2 \\ \vdots \\ \pi_k \end{bmatrix}  
        \ub{
       \begin{bmatrix} 
                 &         &       &         \\
                 &         &       &         \\
         \beta_1 & \beta_2 & \dots & \beta_k \\
                 &         &       &         \\
                 &         &       &         
               \end{bmatrix}}_{B} $
      };
    %\node[anchor=west,right=0.1cm of params] {%
    %\includegraphics[width=4cm,height=2cm,keepaspectratio]{figures/disc.png}
    %};

@@

  \begin{canvas}
    %\input{spectral-experts.tikz}
    \spectralexpertsdiag
    \node[style=txt] at (reg-label) {regression};
    \node[style=txt] at (tf-label) {tensor factorization};

    \point{reg-assumptions}{($(reg-label) - (0,1.0cm)$)};
    \node<2->[style=txt] at (data3 |- reg-assumptions) {
      {\bf Assumptions:}
    };
    \node<3->[style=txt] at (reg-assumptions) {
        $\begin{aligned}
          \hat\E[\vvec(x\tp{2})\tp{2}] &\succ 0\\
          \hat\E[\vvec(x\tp{3})\tp{2}] &\succ 0.
        \end{aligned}$
        };
    \node<4->[style=txt] at (reg-assumptions -| tf-label) {
        $\begin{aligned}
          \pi &\succ 0\\
          rank(B) &= k \le d
        \end{aligned}$
        };
  \end{canvas}
\end{frame}

@@

    \begin{align*}
      \robustaltm<-2>{\hat M_2}{\mathmb{\hat M_3}} &= \arg\min_{M} \sum_{(x,y)\in\mathcal{D}} \left( 
      \robustaltm<-2>{y^2 - \innerp{M}{x\tp{2}}}{\mathmb{y^3 - \innerp{M}{x\tp{3}}}}
      - \textrm{bias}\right)^2 
      \uncover<2->{+
      \robustaltm<2>{\mathmb{\ub{\|M\|_{*}}_{\sum_i \sigma_i(M)}}}
        {\|M\|_{*}}
      } \\
    \end{align*}

@@

  \begin{canvas}
    \spectralexpertsdiag
    \node[style=txt] at (reg-label) {\robustalt<1>{\color{blue}\textbf{low-rank}}{low-rank}regression};
    \node[style=txt] at (tf-label) {tensor factorization};

@@

  \begin{canvas}
    % x, y
    \node[anchor=west] (em) at (-6.5cm,-0.0cm) {%
    \includegraphics<1>[width=\textwidth,height=5cm,keepaspectratio]{figures/1833.pdf}
      \includegraphics<2-3>[width=\textwidth,height=5cm,keepaspectratio]{figures/EM-1833.pdf}
      \includegraphics<4>[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral-1833.pdf}
      \includegraphics<5->[width=\textwidth,height=5cm,keepaspectratio]{figures/Spectral+EM-1833.pdf}
      };

@@

%    \draw<3>[latex-latex,DarkGreen,line width=1pt,dashed] ($(mle) + (-1.2cm,0.8cm)$) -- node[above]{$\mathmg{\epsilon}$} ($(mle) + (+1.2cm,0.8cm)$);
    \node<3->[scale=0.7] at (spec) {x};
    \node<3-> at ($(spec) + (0.5cm,0.3cm)$) {$\mathmg{\hat\theta_{\textrm{spec}}}$};
    \draw<3->[dashed,DarkGreen,line width=0.7pt] ($(spec)-(0.5cm,0)$) -- ($(spec)+(3.5cm,0)$);

    \draw<4->[-latex,smooth,line width=1pt,DarkGreen] ($(spec) + (-0.1cm,+0.05cm)$) -- ($(mle) + (-0.30cm,+0.10cm)$);
    \node<4->[scale=0.7] at (mle) {x};
    \node<4->[anchor=west] at ($(mle) + (0.3cm,0)$) {$\mathmb{\hat\theta}_{\textrm{spec + EM}}$};
    \draw<4->[dashed,blue,line width=0.7pt] ($(mle)-(0.4cm,0)$) -- ($(mle)+(0.4cm,0)$);
  \end{tikzpicture}

@@

\begin{frame}
  \frametitle{Conclusions}
  \begin{itemize}
    \item<+-> Consistent estimator for the mixture of linear regressions
    \item<+-> Expose tensor factorization structure through regression.
    \item<+-> {\bf Theory:} Polynomial sample and computational complexity.
    \item<+-> {\bf Experiments:} Method of moment estimates can be a good initialization for EM.
    \item<+-> {\bf Future Work:} How can we handle other discriminative models?
      \begin{itemize}
          \item<+-> Dependencies between $h$ and $x$ (mixture of experts).
          \item<+-> Non-linear link functions (hidden variable logistic regression).
      \end{itemize}
  \end{itemize}
\end{frame}

@@
